Importance of depth refugia for reef resilience and intervention on the Great Barrier Reef under future climate change

### 2025-10-15 paper/paper.qmd
Update with additional comments from Sam

Beyond individual survival, reef-scale resilience in part depends on connectivity networks
that enable recolonization after disturbances [@angeler2023; @elmhirst2009; @hock2017].
Larval source populations that export more larvae than they import are essential for
system-wide recovery, providing larvae that enable recolonization at distant locations
[@hock2017; @thomas2015]. Yet these source reefs are paradoxically more vulnerable
individually, relying heavily on local recruitment for their own recovery. When disturbances
impact major source populations, cascading effects reduce larval export and slow recovery
throughout the connected network [@figueiredo2022; @hock2017; @munday2009].

@@

Recent advances in whole-ecosystem larval connectivity modelling [@choukroun2025] and coral
thermal response research [@bairos-novak2021] enable integrated investigation of how
multiple factors interact to determine climate resilience. By incorporating the latest coral
population models [@ribeirodealmeida2024], ecosystem-wide connectivity data, and updated
climate projections [@dixon2022], we can examine which factors most strongly influence reef
persistence under different scenarios. While connectivity has been emphasized as critical
for GBR resilience [@hock2017], and remains an important consideration, integrated modelling
reveals depth plays a fundamental role for maintenance of reef coral populations and
positive carbonate budgets under current projections of climate change.

>>>>

Beyond individual survival, reef-scale resilience in part depends on connectivity networks
that enable recolonization after disturbances [@angeler2023; @elmhirst2009; @hock2017].
Larval source populations that have disproportionately numerous and/or strong connections to
other reefs are essential for system-wide recovery, providing larvae that enable recovery at
other reefs after disturbance [@hock2017; @thomas2015]. When disturbances impact major
source populations, cascading effects reduce larval export and slow recovery throughout the
connected network [@figueiredo2022; @hock2017; @munday2009]. Identifying and protecting
source reefs has understandably been a major focus of research and management in recent
years [@greatba2017].

@@

Recent advances in whole-ecosystem larval connectivity modelling [@choukroun2025] and coral
thermal response research [@bairos-novak2021] enable integrated investigation of how
multiple factors interact to determine climate resilience. By incorporating the latest coral
population models [@ribeirodealmeida2024], ecosystem-wide connectivity data, and updated
climate projections [@dixon2022], we can examine which factors most strongly influence reef
persistence under different scenarios. While connectivity has been emphasized as critical
for GBR resilience [@hock2017], and remains an important consideration, our integrated
modelling approach reveals that depth emerges as a dominant predictor for maintenance of
reef coral populations and positive carbonate budgets under current projections of climate
change.
